Chord: A Versatile Platform for Program Analysis

### 2011-04-12 main/doc/properties.tex
updated documentation

\chapter{Chord Properties}
\label{chap:properties}

The only way to specify inputs to Chord is by means of system properties.
There is no command-line argument processing in Chord and any command-line arguments are ignored.
Section \ref{sec:properties-setting} explains how to set properties and
Section \ref{sec:properties-meaning} explains the meaning of properties recognized by Chord.

@@

A property can be passed to Chord in any of several ways.
The reason for providing multiple ways is to provide users shorthand ways for defining properties
once and for all for a particular input Java program, or even once and for all across all Chord runs. 
The following are the different ways by which a property can be passed to Chord in decreasing order of precedence:

@@

On the command-line via the \code{-D<key>=<value>} format.

Use this option to specify properties specific to the current run of Chord.

Typical usage of this option is by running the following command in Chord's \code{main/} directory:
\begin{verbatim}
    prompt> ant -D<key1>=<value1> ... -D<keyN>=<valueN> run
\end{verbatim}

@@

Via the file specified by property \code{chord.props.file}.

Use this option to specify once and for all properties of the program to
be analyzed (e.g., property \code{chord.main.class} specifying the name of that program's main class, property \code{chord.class.path}
specifying the application classpath of that program, etc.).

The default value of this property is \code{[chord.work.dir]/chord.properties} and typically
does not need to be changed by users.
Property \code{chord.work.dir} specifies the directory in which Chord will run.
The default value of this property is the current directory but users must typically set
it on the command line, as follows:

\begin{verbatim}
    prompt> ant -Dchord.work.dir=<...> run
\end{verbatim}
where \code{<...>} denotes the absolute or relative path of the top-level directory of
the program to be analyzed, which contains a file named \code{chord.properties} that users
must create themselves and populate with properties that are recognized by chord and
are specific to that program.

@@

Via file \code{main/chord.properties}.

Use this option to specify once and for all properties you would like to hold in every run of Chord
(e.g., property \code{chord.max.heap} specifying the maximum heap memory size to be used by the
JVM running Chord).
\end{enumerate}

\section{Meaning of Properties}
\label{sec:properties-meaning}

The following properties are recognized by Chord.
The separator for list-valued properties can be either a blank space, a comma, a colon, or a semi-colon.
Notation {\tt [<...>]} is used in this section to denote the value of the property named {\tt <...>}.

@@

\code{chord.props.file}
\begin{quote}
{\bf Type:} location \\
{\bf Description:} Properties file loaded by Chord at the
beginning before doing anything else.  Any of the below properties may
be defined in this file to avoid defining them on the command line
(using the \code{-D<key>=<value>} format) every time Chord is run.
Each relative file/directory name in the value of any property defined
in this file is treated relative to Chord's working directory (which
is specified by property \code{chord.work.dir}). \\
{\bf Default value:} \code{[chord.work.dir]/chord.properties}
\end{quote}

@@

\code{chord.src.path}
\begin{quote}
{\bf Type:} path \\
{\bf Description:} Source path of the input Java program. \\
{\bf Default value:} {\tt ""} \\
{\bf Note:} Chord analyzes only Java bytecode, not Java source code.  This property is used only by the task of converting Java source files into HTML files by analyses that need to present their analysis results at the Java source code level (by calling method \code{chord.program.Program.g().HTMLizeJavaSrcFiles())}.
\end{quote}

@@

\code{chord.reflect.kind}
\begin{quote}
{\bf Type:} {\tt [none|static|dynamic|static\_cast]} \\
{\bf Description:} TODO \\
{\bf Default value:} {\tt none}
\end{quote}

@@

\code{chord.run.analyses}
\begin{quote}
{\bf Type:} string list  \\
{\bf Description:} List of names of analyses to be run in order. \\
{\bf Default value:} {\tt ""} \\
{\bf Note:} If the analysis is written in Java, its name is specified via statement {\tt name=<...>} in its {\tt @Chord} annotation.  If the analysis is written in Datalog, its name is specified via a line of the form ``{\tt \# name=<...>}".
\end{quote}

\code{chord.print.methods}
\begin{quote}
{\bf Type:} string list  \\
{\bf Description:} List of methods whose intermediate representation to print to standard output. \\
{\bf Default value:} {\tt ""} \\
{\bf Note:} Specify each method in format \code{<mname>:<mdesc>@<cname>} where {\tt <mname>} is the method's name, {\tt <mdesc>} is the method's descriptor, and {\tt <cname>} is the name of the method's declaring class. In {\tt <cname>}, use `.' instead of `/', and use {\tt \#} instead of the dollar character. 
\end{quote}

@@

\code{chord.print.rels}
\begin{quote}
{\bf Type:} string list  \\
{\bf Description:} List of names of program relations whose contents must be printed to files \code{[chord.out.dir]/<...>.txt} where {\tt <...>} denotes the relation name. \\
{\bf Default value:} {\tt ""} \\
{\bf Note:} This functionality must be used with caution as certain program relations, albeit represented compactly as BDDs, may contain a large number (e.g., millions) of tuples, resulting in voluminous output when printed in explicit form to a text file.  See Section \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug} target provided in file {\tt build.xml} in Chord's \code{main/} directory).
\end{quote}

@@

\code{chord.classic}
\begin{quote}
{\bf Type:} bool \\
{\bf Description:} TODO \\
{\bf Default value:} TODO
\end{quote}

\code{chord.std.java.analysis.path}
\begin{quote}
{\bf Type:} path \\
{\bf Description:} Partial classpath of analyses written in Java (i.e., {\tt @Chord}-annotated classes).
Conventionally, it includes all such analyses that are predefined in Chord.  \\
{\bf Default value:} The absolute path of file \code{chord.jar}
\end{quote}

\code{chord.ext.java.analysis.path}
\begin{quote}
{\bf Type:} path \\
{\bf Description:} Partial classpath of analyses written in Java (i.e., {\tt @Chord}-annotated classes).
Conventionally, it includes all such external analyses that users want to run. \\
{\bf Default value:} {\tt ""}
\end{quote}

@@

\code{chord.std.dlog.analysis.path}
\begin{quote}
{\bf Type:} path \\
{\bf Description:} Partial classpath of analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}).
Conventionally, it includes all such analyses that are predefined in Chord.  \\
{\bf Default value:} The absolute path of file \code{chord.jar}
\end{quote}

\code{chord.ext.dlog.analysis.path}
\begin{quote}
{\bf Type:} path \\
{\bf Description:} Partial classpath of analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}).
Conventionally, it includes all such external analyses that users want to run. \\
{\bf Default value:} {\tt ""}
\end{quote}

\code{chord.dlog.analysis.path}
\begin{quote}
{\bf Type:} path  \\
{\bf Description:} Complete classpath of analyses written in Datalog (i.e., files with suffix {\tt .dlog} or {\tt .datalog}). \\
{\bf Default value:} \code{[chord.std.dlog.analysis.path]:[chord.ext.dlog.analysis.path]}
\end{quote}

@@

%\item
%\code{chord.save.maps}
% bool
%Write to file \code{[chord.bddbddb.work.dir]/<...>.map} when saving program domain named {\tt <...>}.
%{\tt true}
%{\bf Note:} This functionality is useful for debugging Datalog programs using the {\tt debug} target provided in file {\tt build.xml} in Chord's {\tt main/} directory (see Section \ref{sec:tuning-datalog-analysis}).

>>>>

\xname{properties}
\chapter{Chord Properties}
\label{chap:properties}

The only way to specify inputs to Chord is by means of system properties.  There
is no command-line argument processing in Chord and any command-line arguments
are ignored.  Section \ref{sec:properties-setting} explains how to set
properties and Section \ref{sec:properties-meaning} explains the meaning of
properties recognized by Chord.
Notation {\tt [<...>]} is used in this chapter to denote the value of the property named {\tt <...>}.

@@

A property can be passed to Chord in any of several ways.  The reason for
providing multiple ways is to provide users with shorthand ways for defining
properties once and for all for a particular Java program under analysis, or
even once and for all across all Chord runs.  The following are the different
ways by which a property can be passed to Chord in decreasing order of
precedence:

@@

{\bf How:} On the command-line via the ``\code{-D<key>=<val>}" format.

{\bf When:} Use this option to specify properties specific to the {\it current run} of Chord.

Typical usage of this option is by running the following command:

\begin{framed}
\begin{verbatim}
ant -D<key1>=<val1> ... -D<keyN>=<valN> run
\end{verbatim}
\end{framed}

@@

{\bf How:} Via a user-defined {\it properties file} whose location is specified by property \code{chord.props.file}.

{\bf When:} Use this option to specify once and for all properties of the
Java program to be analyzed (e.g., property \code{chord.main.class} specifying the
name of that program's main class).
Chapter \ref{chap:program-setup} presents an example properties file 
that defines the program properties that are most commonly used.  Section \ref{sec:program-props}
presents all program properties that are recognized by Chord.

The default value of property \code{chord.props.file} is \code{[chord.work.dir]/chord.properties}.
Property \code{chord.work.dir} specifies the directory in which Chord must run; its default value
is the current directory.

There are three ways to use this option of setting properties:

The first way is to override the default value of property \code{chord.work.dir} on the command-line.
This requires naming the above properties file as \code{chord.properties}, placing it in the
directory in which Chord will run, denoted \code{<WORK_DIR>}
(e.g., this could be the top-level directory of the program to be analyzed),
and running the following command:

\begin{framed}
\begin{verbatim}
ant -Dchord.work.dir=<WORK_DIR> run
\end{verbatim}
\end{framed}

The second way is to override the default value of property \code{chord.props.file} on the command-line.
In this case, the properties file can be in any user-desired location, denoted \code{<PROPS_FILE>}, and Chord
will run in the current directory:

\begin{framed}
\begin{verbatim}
ant -Dchord.props.file=<PROPS_FILE> run
\end{verbatim}
\end{framed}

The third (and most flexible) way is to override the default values of both properties \code{chord.work.dir}
and \code{chord.props.file} on the command-line.

@@

{\bf How:} Via the properties file named \code{chord.properties} that is already provided in Chord's main directory.

{\bf When:} Use this option to specify once and for all properties you would
like to hold in {\it every run} of Chord (e.g., property \code{chord.max.heap}
specifying the maximum heap memory size to be used by the JVM running Chord).
\end{enumerate}

\section{Recognized Properties}
\label{sec:properties-meaning}

The following properties are recognized by Chord.
The separator for list-valued properties can be either a blank space, a comma, a colon, or a semi-colon.

@@

\code{chord.props.file}
\begin{quote}
{\bf Type:} location \\
{\bf Description:} Properties file loaded by Chord at the
beginning before doing anything else.  Any of the below properties may
be defined in this file to avoid defining them on the command line
(using the \code{-D<key>=<val>} format) every time Chord is run.
Each relative file/directory name in the value of any property defined
in this file is treated relative to Chord's working directory (which
is specified by property \code{chord.work.dir}). \\
{\bf Default value:} \code{[chord.work.dir]/chord.properties}
\end{quote}

@@

\code{chord.src.path}
\begin{quote}
{\bf Type:} path \\
{\bf Description:} Source path of the input Java program. \\
{\bf Default value:} {\tt ""} \\
{\bf Note:} Chord analyzes only Java bytecode, not Java source code.  This property is used only by the task of converting Java source files into HTML files by analyses that need to present their results at the Java source code level (by calling method \code{chord.program.Program.g().HTMLizeJavaSrcFiles())}.
\end{quote}

@@

\code{chord.reflect.kind}
\begin{quote}
{\bf Type:} {\tt [none|dynamic|static|static\_cast]} \\
{\bf Description:} Algorithm to resolve reflection.  The choices are {\tt none} (do not resolve any reflection),
{\tt dynamic} (run the program and observe how reflection is resolved),
{\tt static} (resolve reflection statically but without analyzing casts), and
{\tt static\_cast} (resolve reflection statically by analyzing casts). \\
{\bf Default value:} {\tt none}
\end{quote}

@@

\code{chord.run.analyses}
\begin{quote}
{\bf Type:} string list  \\
{\bf Description:} List of names of analyses to be run in order. \\
{\bf Default value:} {\tt ""} \\
{\bf Note:} If the analysis is written in Java, its name is specified via statement {\tt name="..."} in its {\tt @Chord} annotation.  If the analysis is written in Datalog, its name is specified via a line of the form ``{\tt \# name=...}".
\end{quote}

\code{chord.print.methods}
\begin{quote}
{\bf Type:} string list  \\
{\bf Description:} List of methods whose intermediate representation to print to standard output. \\
{\bf Default value:} {\tt ""} \\
{\bf Note:} Specify each method in format \code{mname:mdesc@cname} where {\tt mname} is the method's name, {\tt mdesc} is the method's descriptor, and {\tt cname} is the name of the method's declaring class. In {\tt cname}, use `.' instead of `/', and use `{\tt \#}' instead of `{\tt \$}'. 
\end{quote}

@@

\code{chord.print.rels}
\begin{quote}
{\bf Type:} string list  \\
{\bf Description:} List of names of program relations whose contents must be printed to files \code{[chord.out.dir]/<RELNAME>.txt} where
{\tt <RELNAME>} denotes the relation name. \\
{\bf Default value:} {\tt ""} \\
{\bf Note:} This functionality must be used with caution as certain program relations, albeit represented compactly as BDDs, may contain
 a large number (e.g., millions) of tuples, resulting in voluminous output when printed in explicit form to a text file.  See Section
 \ref{sec:tuning-datalog-analysis} for a more efficient way to query the contents of program relations (namely, by using the {\tt debug}
 target provided in file {\tt build.xml} in Chord's main directory).
\end{quote}

@@

\code{chord.classic}
\begin{quote}
{\bf Type:} bool \\
{\bf Description:} Whether to use the classic project (as opposed to the modern project).  See Chapter\ref{chap:analyses} for the difference
 between the two kinds of projects. \\
{\bf Default value:} \code{true}
\end{quote}

\code{chord.std.java.analysis.path}
\begin{quote}
{\bf Type:} path \\
{\bf Description:} Partial classpath of analyses written in Java (i.e., {\tt @Chord}-annotated classes).
Conventionally, it includes all Java analyses that are predefined in Chord.  \\
{\bf Default value:} The absolute path of file \code{chord.jar}.
\end{quote}

\code{chord.ext.java.analysis.path}
\begin{quote}
{\bf Type:} path \\
{\bf Description:} Partial classpath of analyses written in Java (i.e., {\tt @Chord}-annotated classes).
Conventionally, it includes all user-defined analyses. \\
{\bf Default value:} {\tt ""}
\end{quote}

@@

\code{chord.std.dlog.analysis.path}
\begin{quote}
{\bf Type:} path \\
{\bf Description:} Partial path of analyses written in Datalog (i.e., files with suffix {\tt .dlog}).
Conventionally, it includes all Datalog analyses that are predefined in Chord. \\
{\bf Default value:} The absolute path of file \code{chord.jar}.
\end{quote}

\code{chord.ext.dlog.analysis.path}
\begin{quote}
{\bf Type:} path \\
{\bf Description:} Partial path of analyses written in Datalog (i.e., files with suffix {\tt .dlog}).
Conventionally, it includes all user-defined Datalog analyses. \\
{\bf Default value:} {\tt ""}
\end{quote}

\code{chord.dlog.analysis.path}
\begin{quote}
{\bf Type:} path  \\
{\bf Description:} Complete path of analyses written in Datalog (i.e., files with suffix {\tt .dlog}). \\
{\bf Default value:} \code{[chord.std.dlog.analysis.path]:[chord.ext.dlog.analysis.path]}
\end{quote}

@@

%\item
%\code{chord.save.maps}
% bool
%Write to file \code{[chord.bddbddb.work.dir]/<...>.map} when saving program domain named {\tt <...>}.
%{\tt true}
%{\bf Note:} This functionality is useful for debugging Datalog programs using the {\tt debug} target provided in file {\tt build.xml} in Chord's main directory (see Section \ref{sec:tuning-datalog-analysis}).

### 2011-04-12 main/doc/writing_analysis.tex
updated documentation

\section{Writing a Program Analysis}
\label{sec:writing-analysis}

TODO

>>>>

\xname{writing_analysis}
\chapter{Writing an Analysis}
\label{chap:writing-analysis}

Chord provides several {\it analysis templates}: classes containing boilerplate
code that can be extended by users to rapidly prototype different kinds of
analyses.  These templates are organized in the following hierarchy in
package \javadoc{chord.project.analyses}{chord/project/analyses/package-summary.html}:

\begin{verbatim}
JavaAnalysis
    |
    |--- ProgramDom
    |
    |--- ProgramRel
    |
    |--- DlogAnalysis
    |
    |--- RHSAnalysis
    |        |
    |        |--- ForwardRHSAnalysis
    |        |
    |        |--- BackwardRHSAnalysis
    |
    |--- BasicDynamicAnalysis
             |
             |--- DynamicAnalysis
\end{verbatim}

All these classes implement
interface \javadoc{chord.project.ITask}{chord/project/ITask.html} and can
therefore serve as tasks.  We next describe each of them in more detail.

\section{JavaAnalysis}
\label{sec:java-analysis}

Class \code{chord.project.analyses.JavaAnalysis} provides the most general
template for writing an analysis imperatively in Java.  An analysis can be
created using this template by extending this class as follows:

\begin{verbatim}
@Chord(
    name = "example-java",
    consumes = { "C1", ..., "Cn" },
    produces = { "P1", ..., "Pm" },
    types = { A1.class, ..., An.class, B1.class, ..., Bm.class },
    namesOfTypes = { "C1", ..., "Cn", "P1", ..., "Pm" },
)
public class ExampleAnalysis extends JavaAnalysis {
    @Override
    public void run() {
        ClassicProject project = ClassicProject.g();

        A1 c1 = (A1) project.getTrgt("C1");
        ...
        An cn = (An) project.getTrgt("Cn");

        B1 p1 = (B1) project.getTrgt("P1");
        ...
        Bm pm = (Bm) project.getTrgt("Pm");

        ... // analysis code computing p1, ..., pm from c1, ..., cn
    }
}
\end{verbatim}

The class must be compiled to a {\tt .class} file that occurs in
some element (directory or jar/zip file) of the path
specified by property \code{chord.java.analysis.path}.  Chord represents the
corresponding analysis using a unique object of the class.

The class is annotated with an {\tt @Chord} annotation defined in
class \javadoc{chord.project.Chord}{chord/project/Chord.html}.  This annotation
allows specifying the name of the analysis ({\tt example-java}), the names of
the targets that are directly consumed ({\tt C1}, ..., {\tt Cn}) and directly
produced ({\tt P1}, ..., {\tt Pm}) by the analysis, and the types of targets.
There is a one-to-one correspondence between the arrays denoted by fields {\tt
types} and {\tt namesOfTypes}, e.g., the type of the target named {\tt C1} is
class {\tt A1}, and so on.  In principle, the type of {\it any} target in the
project can be specified here.  In practice, however, the types of only the
targets declared as directly consumed/produced by this analysis are specified.
Moreover, although the type of each target that is directly consumed/produced by
the above analysis is specified in the above annotation, in practice the types
of hardly any targets need to be explicitly specified, because they can be
automatically inferred by Chord from analyses created using more specialized
templates discussed below that also consume/produce those targets.

The code of the analysis must be supplied in the {\tt run()} method.  This
method typically retrieves the representation of each directly consumed/produced
target from the project, reads (the state of) the consumed targets, performs
some computation over it, and writes to (the state of) the produced targets.

The analysis templates presented in the following sections are more specialized
forms of the JavaAnalysis template in that they restrict what the directly
consumed/produced targets and the analysis code in the {\tt run()} method can
be.

%called JavaAnalysis (Section \ref{sec:java-analysis}), and the other for writing
%analyses declaratively in Datalog, called DlogAnalysis
%(Section \ref{sec:dlog-analysis}).  JavaAnalysis is the most general template:
%it least constrains the kind of analysis that can be written but also offers the
%least boilerplate code for writing an analysis.  Chord provides several
%additional templates for writing analyses imperatively in Java that are
%specialized forms of JavaAnalysis, namely, ProgramDom
%(Section \ref{sec:program-dom}), ProgramRel (Section \ref{sec:program-rel}),
%RHSAnalysis (Section \ref{sec:rhs-analysis}), and DynamicAnalysis
%(Section \ref{sec:dynamic-analysis}).  We next explain each of these templates.

\section{DlogAnalysis}
\label{sec:dlog-analysis}

A common way to rapidly prototype analyses in Chord is using a declarative
logic-programming language called Datalog.  A Datalog analysis can be defined in
any {\tt .dlog} file that occurs in some element (directory or jar/zip file)
of the path specified by property \code{chord.dlog.analysis.path}.  Chord represents each such analysis
using a unique object of class \javadoc{chord.project.analyses.DlogAnalysis}{chord/project/analyses/DlogAnalysis.html}.

A Datalog analysis declares a bunch of input/output {\it program relations},
each over one or more {\it program domains}, and provides a bunch of rules
(constraints) specifying how to compute the output relations from the input
relations.

TODO: explain inputs, outputs, code

Briefly, each Datalog analysis specifies the following:
\begin{itemize}
\item
a set of program domains D1, ..., Dk; a program domain is a collection of program artifacts of a particular kind.
\item
a set of input program relations I1, .., In, each over one of more program domains D1, ..., Dk;
a program relation
\item
I1, ..., In, each over one or more program domains D1, ..., Dk
, of which , a set of output program
relations O1, ..., Om, a set of rules specifying how to compute the output relations from the input
relations, and a set of program domains D1, ..., Dk.  
\end{itemize}

Chapter \ref{chap:datalog-analysis} explains how to write a Datalog analysis in
more detail.

\section{ProgramDom}
\label{sec:program-dom}

\section{ProgramRel}
\label{sec:program-rel}

\section{DynamicAnalysis}
\label{sec:dynamic-analysis}

\section{RHSAnalysis}
\label{sec:rhs-analysis}

%\begin{itemize}
%\item
%B is the domain of basic blocks.
%\item
%C is the domain of abstract calling contexts.
%\item
%E is the domain of program points that read or write a field.
%\item
%F is the domain of fields inside a class.
%\item
%H is the domain of allocation sites.
%\item
%I is the domain of invocation statements (method calls)
%\item
%L is the domain of lock acquisition points.
%\item
%M is the domain of program methods.
%\item
%P is the domain of simple statements. %SAY MORE HERE?
%\item
%R is the domain of lock release points.
%\item
%T is the domain of types (classes and primitives).
%\item
%V is the domain of reference-typed variables.
%\item
%W is the domain of loop head statements. %SAY MORE HERE?
%\item
%Z is a domain of integers corresponding to argument positions.The size of the domain is chosen to 
%accommodate the method with the greatest number of arguments.
%\end{itemize}

%\section{Core relations}

%Chord includes code to build a very large number of relations. These standard relations are very useful in writing your own analyses.  For example, relation \texttt{sub} includes all pairs of types \texttt{(t1, t2)} where \texttt{t2} is a subclass of \texttt{t1}.  Most of these standard relations are in the package \texttt{chord.rels}.

%A Java analysis should extend JavaAnalysis or ProgramRel.  It's usually easiest if an analysis only produces one relation, in which case extending ProgramRel is the best approach.  If you extend ProgramRel, all the work of creating and saving the relation is taken care of for you; all you need to do is add tuples to the relation.Typically you do this by calling \texttt{super.add(...)} with the tuple you wish to add to the relation.  For forther convenience, you can implement one of the Visitor interfaces defined in \texttt{chord.program.visitors}.  This will force you to implement some \texttt{visit} methods that iterate over all elements in some domain, such as all methods or all invocation points.

%Sometimes, you need to create two different relations at the same time. In this case, the right approach is to extend JavaAnalysis, and explicitly create and save the relations on your own.

%Regardless of which of these approaches you take, every Java analysis must have an \texttt{@Chord} annotation.  This annotation is a set of named fields.  At a minimum, this annotation must specify the name of the analysis.  If you extend ProgramRel, the name of the analysis will be treated as the same as that of the generated relation.  For example, \texttt{@Chord(name = "F")} is a valid annotation.  The name of the Java class that generates the relation is ignored by Chord -- only the annotation matters.

%\textbf{MAYUR, IS SIGN MANDATORY?   IF NOT, WHEN CAN IT BE OMITTED?}

%There are also other annotation fields available.  You may specify the dependencies of your analysis via \texttt{consumes = \{ "xyz"\}}.

%If your analysis does not extend ProgramRel, you should also specify the relations that your analysis produces. You do this via the \texttt{produces} relation.

%Should mention the signs fields.
